Quantum Walks and Application to Quantum Money\\[1em] \normalsize McMaster University\\ Department of Computing and Software

### 2025-04-12 draft.tex
draft updated

In 2009, Aaronson~\cite{Aaronson2009} proposed the first \emph{public-key quantum money} scheme, in which anyone can verify a quantum banknote while only the bank can mint (issue) new valid banknotes. Public-key quantum money removes the transactional bottleneck of involving the bank every time. Aaronson’s specific scheme was later broken by Lutomirski et al.~\cite{Lutomirski2010}. In the years since, several alternative public-key quantum money constructions have been explored. However, each of these proposals has either been broken by further cryptanalysis or relies on unconventional cryptographic assumptions, making their security or practicality questionable.

@@

The resolution is to enlarge the Hilbert space by introducing an auxiliary \emph{coin register} that provides additional degrees of freedom for unitary evolution. This model, first proposed by Watrous in the context of quantum graph algorithms, leads to the \emph{coined discrete-time quantum walk}.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Invariant Subspaces and Symmetry Reduction}

@@

\subsection{Hitting and Mixing Properties}

Quantum walks on Cayley graphs have been applied to study hitting and mixing times. In the classical case, random walks on many Cayley graphs are rapidly mixing (e.g., those based on nonabelian simple groups, which often form expanders).

In the quantum case, mixing is more subtle due to unitary evolution. One typically considers \emph{instantaneous} or \emph{time-averaged} mixing. Some Cayley graphs support perfect state transfer or periodicity in CTQWs.

\textbf{Examples:}
\begin{itemize}
    \item On the cycle graph $C_n$, a CTQW exhibits perfect revival at time $t = \pi$ if $n$ is even.
    \item On the hypercube, the CTQW exhibits approximate uniform mixing at specific times.
\end{itemize}

These properties are derived from the spectral structure, which is accessible via group characters.

\subsection{Discrete-Time Walks and Symmetry}

Coined DTQWs also benefit from Cayley graph symmetry. If a symmetric coin (e.g., the Grover coin) is used at each vertex of a vertex-transitive graph, the walk respects the full symmetry. As a result:
\begin{itemize}
    \item Amplitudes often remain uniform on symmetry-equivalent vertices.
    \item Walk dynamics can often be analyzed by reducing to a lower-dimensional invariant subspace.
    \item In some cases, symmetry can lead to \emph{localization}, where amplitude becomes trapped due to spectral properties of the coin.
\end{itemize}

These effects can be mitigated or exploited through careful coin design or by using Szegedy's method, which naturally respects symmetry.

\subsection{Summary}

Cayley graphs offer a rich and powerful structure for quantum walks. Their group-theoretic properties allow:
\begin{itemize}
    \item Spectral analysis via irreducible representations.
    \item Dimensionality reduction using symmetry and invariant subspaces.
    \item Explicit calculation of probabilities, mixing behavior, and hitting times.
\end{itemize}

Many efficient quantum walk algorithms exploit these properties either directly or implicitly. In subsequent sections, we will explore quantum walks on more general structures where group actions go beyond simple Cayley graphs.

@@

We require two properties from such a scheme:
\begin{description}
    \item[Correctness:] For any valid output $(s, |\$\rangle)$ of $\mathsf{Mint}(1^n)$, the verification algorithm accepts with high probability. Formally, $\Pr[\mathsf{Verify}(s, |\$\rangle) = \mathtt{accept}]$ should be very close to 1 (typically we require it to be $1 - negl(n)$, where $negl(n)$ denotes a negligible function in $n$).
    \item[Security (Unforgeability):] No efficient adversary, given one or more genuine banknotes, can produce a new valid banknote. In the strongest formulation, if an adversary is given a single valid $(s, |\$\rangle)$, it should be infeasible for them to produce two quantum states $|\$'\rangle$ and $|\$''\rangle$ such that $\mathsf{Verify}(s, |\$'\rangle) = \mathsf{Verify}(s, |\$''\rangle) = \mathtt{accept}$ (i.e., they cannot create a second copy that passes verification while retaining the original). More generally, even given multiple banknotes, the adversary should not be able to create new ones with fresh serial numbers that pass verification. This captures the idea that quantum money cannot be counterfeited.
\end{description}

In a \emph{public-key} scheme, the verifier uses only the public information (here, the serial number and any scheme-wide public parameters) to check validity, and does not need any secret key. The bank (mint) may have a secret key to produce the states, but after minting, the bank’s involvement is not needed. In our scheme, the serial number will effectively contain the information needed for public verification (similar to how a classical bill might have a number anyone can look up, except here it’s used in a quantum procedure).

@@

\section{Fourier-Based Quantum Money from Group Actions}
Zhandry’s quantum money scheme \cite{Zhandry2024} is built on an abelian group action $G \curvearrowright X$. We have a family of group actions parameterized by the security parameter $n$ (so the size of $G$ grows with $n$). For simplicity, assume $|G| = N$. The scheme also fixes a particular element $x_0 \in X$ that everyone knows (a public base point in $X$).

The minting algorithm $\mathsf{Mint}$ in Zhandry’s scheme works as follows (informally):
\begin{enumerate}
    \item On input $1^n$, sample a random group element $h \leftarrow G$. This $h$ will serve as the serial number $s$ for the banknote.
    \item Prepare two registers: the first is a group element register, initialized to $|0_G\rangle$ (an encoding of the identity element of $G$), and the second is an $X$-register initialized to $|x_0\rangle$.
    \item Apply the quantum Fourier transform over $G$ to the first register. This creates a uniform superposition over $G$: 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |x_0\rangle\,.
    \] 
    \item Now apply the group action in a controlled manner: for each basis state $|g\rangle$ in the first register, apply $g$’s action on the second register. This transforms the joint state into 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    which is an entangled state between a superposition of group elements and their corresponding $X$-elements.
    \item Apply the inverse QFT (which is the adjoint of the QFT) on the first register. This recombines the amplitudes in the first register. The result (by properties of the QFT) can be shown to be 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} \omega^{\langle h,\,g \rangle} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    where $\omega^{\langle h,\,g \rangle}$ denotes the phase picked up that depends on $h$ and $g$. In fact, this state can be written as $|h\rangle \otimes \frac{1}{\sqrt{N}}\sum_{g \in G} |g * x_0\rangle$; essentially, the first register ends up in $|h\rangle$ (the serial number) and the second register is an equal superposition of all elements of $X$ in the orbit of $x_0$ (which is $X$ itself, since the action is transitive) but weighted by a phase related to $h$. This is the so-called group-action Fourier state.
    \item The first register (which contains $|h\rangle$) is measured in the computational basis, yielding the random serial number $g$ and collapsing the second register to a state that we will denote $|\$_h\rangle$. This $|\$_h\rangle$ is the actual quantum money state corresponding to serial $h$.
    \item Output the pair $(s = h,\; |\$\rangle = |\$_h\rangle)$.
\end{enumerate}

@@

\section{Hartley-Transform Money Scheme Construction}
Our scheme modifies the above Fourier-based construction by utilizing the \textbf{quantum Hartley transform (QHT)} in place of the QFT. Since the Hartley transform is closely related to the Fourier transform (essentially it produces the real and imaginary parts in a single real-valued transform), one might expect this substitution to be straightforward. And indeed, for the minting procedure it largely is: we will use the QHT to create real-amplitude superposition states.

Concretely, assume again a group action $G \curvearrowright X$ with $|G|=N$. We fix the same generating set $S$ and basepoint $x_0$. The $\mathsf{Gen}_{H}$ (Hartley generation) algorithm:
\begin{enumerate}
    \item Sample a random $h \leftarrow G$ to be the serial number.
    \item Prepare $|0_G, x_0\rangle$ as before.
    \item Apply the \emph{quantum Hartley transform} over $G$ to the first register. The Hartley transform (on an abelian group of order $N$) is defined as:
    \[
       \mathrm{QHT}: |g\rangle \mapsto \frac{1}{\sqrt{N}} \sum_{k \in G} \cas \Big(\frac{2\pi \langle k,g \rangle}{N}\Big) |k\rangle,
    \] 
    where $\mathrm{cas}(\theta) = \cos \theta + \sin \theta$ and $\langle k,g \rangle$ denotes some pairing like an exponent or dot product (for cyclic groups this is just $kg$). In simpler terms, $\mathrm{QHT}(|0_G\rangle) = \frac{1}{\sqrt{N}}\sum_{k \in G} |k\rangle$ (because $\cas(0)=1$), \emph{the same uniform superposition as the QFT produces for $|0\rangle$}. So after this step, we again have $\frac{1}{\sqrt{N}}\sum_{g \in G} |g\rangle|x_0\rangle$, but crucially the coefficients are all \emph{real} and positive (in fact all $1/\sqrt{N}$).
    \item Perform the controlled group action by the first register on the second: $\frac{1}{\sqrt{N}}\sum_{g} |g\rangle |g*x_0\rangle$.
    \item Apply the inverse QHT on the first register (which being its own inverse up to a constant factor, is the same operation again since Hartley is symmetric like Fourier). Now here is where things differ: after this operation, the joint state will be 
    \[
       \frac{1}{\sqrt{N}} \sum_{g \in G} \cas\Big(\frac{2\pi \langle h,g \rangle}{N}\Big)\, |g\rangle \otimes |g*x_0\rangle.
    \] 
    Because the Hartley transform combines what would be cosine and sine terms, this state is not as nicely factorable as in the Fourier case. In the Fourier case, we got $|h\rangle \otimes \text{(orbit state)}$. In the Hartley case, we do not get a single basis state in the first register; instead we get a superposition in the first register with real coefficients $c_g = \cas(2\pi\langle h,g\rangle/N)$. However, note that $c_g$ is just some real number in $[-1,1]$ depending on $h$ and $g$. Another way to express the above state is:
    \[
      |\$\rangle = \frac{1}{\sqrt{N}} \sum_{x \in X} \Big(\frac{1}{\sqrt{N}}\sum_{g: g*x_0 = x} \cas\big(\frac{2\pi \langle h,g \rangle}{N}\big)\Big)\, |x\rangle.
    \] 
    Since for a given $x\in X$, all $g$ such that $g*x_0=x$ have the form $g = u h$ for a unique $u \in G$ (because the action is free and transitive), one can simplify the internal sum using properties of the cas function. But we need not dive into that here.
    \item No measurement of the first register is performed. In fact, we do not have a nice product state separation, so we cannot simply measure one register to get $h$. Instead, the output of minting is just the quantum state $|\$\rangle$ (which is entangled between what used to be two registers) and the known classical serial number $h$ (we know what $h$ we picked).
\end{enumerate}

@@

% \begin{algorithm}[$\qht_N$] \
\paragraph{Algorithm $\qht_N$}
    % \label{alg:qht-N}
    % \begin{description}[font = \normalfont\itshape, itemsep = 0mm, parsep = 0mm, topsep = 1mm]
    %     \item [Input:] quantum state $\ket{\psi} \in \C^N$, where $N = 2^n$
    %     \item [Output:] quantum state $\qht_N\ket{\psi}$
    % \end{description}

@@

% \begin{theorem}
\paragraph{Theorem.}
    \label{thm:qht-cost}
    Algorithm above is correct and can be implemented using $\approx \log^2 N + O(\log N)$ elementary gates.
% \end{theorem}

% \begin{proof}
\paragraph{Proof.}
    The correctness of the algorithm follows from the preceding discussion. Except for the unitary \( U_R \) and the negation unitary in Step~\ref{stp:ngt}, all steps of the algorithm can be implemented using \( O(1) \) elementary gates. The negation operation in Step~\ref{stp:ngt} can be realized using approximately \( \lceil \log N \rceil \) elementary gates.

@@

    Let \( T(N) \) denote the gate complexity of the algorithm for an input of dimension \( N \). Assuming access to the \( R_j \) gates as elementary operations, the recurrence relation becomes
    \[
    T(N) \approx T(N / 2) + 2\log N + O(1).
    \]
    Solving this recurrence yields \( T(N) \approx \log^2 N + O(\log N) \), as claimed.
%     \end{proof}
%%%%%%%%%%%  ta inja  %%%%%%%%%%%

@@

For the purposes of this thesis, the main point is: \emph{we can implement the Hartley transform needed for our money scheme with polynomial efficiency}. Thus, switching to the Hartley transform does not introduce any prohibitive cost. All steps of minting and verification remain efficient. Having covered the construction of the money scheme and the necessary tools, we now proceed to the final piece: the detailed verification algorithm using quantum walks, and an analysis of how it validates genuine banknotes and foils counterfeit attempts.
\vspace{1em}

@@

    % \begin{lemma}
    \paragraph{Lemma.}
        \label{lem:h-eigen}
        The money state $\ket{\mathbb{Z}_N^{(h)} * x}_H$ is an eigenstate of $W$ with eigenvalue $e^{i\lambda_h t}$.
    % \end{lemma}
    % \begin{proof}
    \paragraph{Proof.}
        We have 
        \begin{align*}
            e^{iAt} \ket{\mathbb{Z}_N^{(h)} * x}_H
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x} \braket{\mathbb{Z}_N^{(g)} * x}{\mathbb{Z}_N^{(h)} * x}_H \\
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x} \bra{\mathbb{Z}_N^{(g)} * x} \Big( \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} \ket{\mathbb{Z}_N^{(-h)} * x} \Big) \\
            & = e^{i\lambda_{h}t} \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} e^{i\lambda_{-h}t} \ket{\mathbb{Z}_N^{(-h)} * x} \\
            & = e^{i\lambda_{h}t} \ket{\mathbb{Z}_N^{(h)} * x}_H,
        \end{align*}
        where the second equality follows from the identity in \eqref{eq:QHT_QFT}, and the last equality follows from the fact that $\lambda_h = \lambda_{-h}$.
    % \end{proof}

>>>>

In 2009, Aaronson~\cite{Aaronson2009} proposed the first \emph{public-key quantum money} scheme, in which anyone can verify a quantum banknote while only the bank can generate new valid banknotes. Public-key quantum money removes the transactional bottleneck of involving the bank every time. Aaronson’s specific scheme was later broken by Lutomirski et al.~\cite{Lutomirski2010}. In the years since, several alternative public-key quantum money constructions have been explored. However, each of these proposals has either been broken by further cryptanalysis or relies on unconventional cryptographic assumptions, making their security or practicality questionable.

@@

The resolution is to enlarge the Hilbert space by introducing an auxiliary \emph{coin register} that provides additional degrees of freedom for unitary evolution. This model, first proposed by Watrous \cite{Watrous2001} and later formalized in algorithmic contexts by Aharonov et al.~\cite{Aharonov2001}, leads to the \emph{coined discrete-time quantum walk}.

@@

Coined quantum walks on simple structures such as the line were also analyzed in early work by Nayak and Vishwanath~\cite{Nayak2000}, who studied their spreading properties and asymptotic behavior.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Structure and Symmetry in Cayley Graph Walks}

@@

% \paragraph{Hitting and Mixing Properties}

% Quantum walks on Cayley graphs have been applied to study hitting and mixing times. In the classical case, random walks on many Cayley graphs are rapidly mixing (e.g., those based on nonabelian simple groups, which often form expanders).

% In the quantum case, mixing is more subtle due to unitary evolution. One typically considers \emph{instantaneous} or \emph{time-averaged} mixing. Some Cayley graphs support perfect state transfer or periodicity in CTQWs.

% \textbf{Examples:}
% \begin{itemize}
%     \item On the cycle graph $C_n$, a CTQW exhibits perfect revival at time $t = \pi$ if $n$ is even.
%     \item On the hypercube, the CTQW exhibits approximate uniform mixing at specific times.
% \end{itemize}

% These properties are derived from the spectral structure, which is accessible via group characters.

% \paragraph{Discrete-Time Walks and Symmetry}

% Coined DTQWs also benefit from Cayley graph symmetry. If a symmetric coin (e.g., the Grover coin) is used at each vertex of a vertex-transitive graph, the walk respects the full symmetry. As a result:
% \begin{itemize}
%     \item Amplitudes often remain uniform on symmetry-equivalent vertices.
%     \item The dynamics of the walk can often be simplified by restricting to a smaller symmetric subspace.
%     \item In some cases, symmetry can lead to \emph{localization}, where amplitude becomes trapped due to spectral properties of the coin.
% \end{itemize}

% These effects can be mitigated or exploited through careful coin design or by using Szegedy's method, which naturally respects symmetry.

% In summary, Cayley graphs provide a rich and algebraically structured setting for quantum walks. Their underlying group-theoretic properties enable powerful techniques such as spectral analysis through irreducible representations, dimensionality reduction via symmetries and invariant subspaces, and explicit calculations of probabilities, mixing behavior, and hitting times. These features make Cayley graphs especially amenable to efficient quantum walk algorithms, which often exploit such structures either explicitly or implicitly. In the sections that follow, we extend these ideas to more general settings where group actions go beyond the framework of simple Cayley graphs.

\paragraph{Hitting and Mixing Properties}
Quantum walks on Cayley graphs have been applied to study hitting and mixing times. In the classical case, random walks on many Cayley graphs are rapidly mixing (e.g., those based on nonabelian simple groups, which often form expanders).

In the quantum case, mixing is more subtle due to unitary evolution. One typically considers instantaneous or time-averaged mixing. Some Cayley graphs support perfect state transfer or periodicity in CTQWs~\cite{Gerhardt2003}. Recent work has also established upper bounds on the mixing time for quantum walks on random graphs, showing polynomial-time convergence in certain regimes~\cite{Chakraborty2020}.

Examples:
\begin{itemize}
    \item On the cycle graph $C_n$, a CTQW exhibits perfect revival at time $t = \pi$ if $n$ is even.
    \item On the hypercube, the CTQW exhibits approximate uniform mixing at specific times.
\end{itemize}
These properties are derived from the spectral structure, which is accessible via group characters.

\paragraph{Discrete-Time Walks and Symmetry}
Coined DTQWs also benefit from Cayley graph symmetry. If a symmetric coin (e.g., the Grover coin) is used at each vertex of a vertex-transitive graph, the walk respects the full symmetry. As a result:
\begin{itemize}
    \item Amplitudes often remain uniform on symmetry-equivalent vertices.
    \item The dynamics of the walk can often be simplified by restricting to a smaller symmetric subspace.
    \item In some cases, symmetry can lead to localization, where amplitude becomes trapped due to spectral properties of the coin.
\end{itemize}
These effects can be mitigated or exploited through careful coin design or by using Szegedy’s method, which naturally respects symmetry.

In summary, Cayley graphs provide a rich and algebraically structured setting for quantum walks. Their underlying group-theoretic properties enable powerful techniques such as spectral analysis through irreducible representations, dimensionality reduction via symmetries and invariant subspaces, and explicit calculations of probabilities, mixing behavior, and hitting times. These features make Cayley graphs especially amenable to efficient quantum walk algorithms, which often exploit such structures either explicitly or implicitly. In the sections that follow, we extend these ideas to more general settings where group actions go beyond the framework of simple Cayley graphs.

@@

We require two properties from such a scheme:
\begin{description}
    \item[Correctness:] For any valid output $(s, |\$\rangle)$ of $\mathsf{Gen}(1^n)$, the verification algorithm accepts with high probability. Formally, $\Pr[\mathsf{Verify}(s, |\$\rangle) = \mathtt{accept}]$ should be very close to 1 (typically we require it to be $1 - \negl(n)$, where $\negl(n)$ denotes a negligible function in $n$).
    \item[Security (Unforgeability):] No efficient adversary, given one or more genuine banknotes, can produce a new valid banknote. In the strongest formulation, if an adversary is given a single valid $(s, |\$\rangle)$, it should be infeasible for them to produce two quantum states $|\$'\rangle$ and $|\$''\rangle$ such that $\mathsf{Verify}(s, |\$'\rangle) = \mathsf{Verify}(s, |\$''\rangle) = \mathtt{accept}$ (i.e., they cannot create a second copy that passes verification while retaining the original). More generally, even given multiple banknotes, the adversary should not be able to create new ones with fresh serial numbers that pass verification. This captures the idea that quantum money cannot be counterfeited.
\end{description}

In a \emph{public-key} scheme, the verifier uses only the public information (here, the serial number and any scheme-wide public parameters) to check validity, and does not need any secret key. The bank may have a secret key to produce the states, but after generation, the bank’s involvement is not needed. In our scheme, the serial number will effectively contain the information needed for public verification (similar to how a classical bill might have a number anyone can look up, except here it’s used in a quantum procedure).

@@

\section{Fourier-Based Quantum Money from Group Actions}
Zhandry’s quantum money scheme \cite{Zhandry2024} is built on an abelian group action $*:G \times X \mapsto X$. We have a family of group actions parameterized by the security parameter $n$ (so the size of $G$ grows with $n$). For simplicity, assume $|G| = N$. The scheme also fixes a particular element $x_0 \in X$ that everyone knows (a public base point in $X$).

The generation algorithm $\mathsf{Gen}$ in Zhandry’s scheme works as follows (informally):
\begin{enumerate}
    \item On input $1^n$, sample a random group element $h \leftarrow G$. This $h$ will serve as the serial number $s$ for the banknote.
    \item Prepare two registers: the first is a group element register, initialized to $|0_G\rangle$ (an encoding of the identity element of $G$), and the second is an $X$-register initialized to $|x_0\rangle$.
    \item Apply the quantum Fourier transform over $G$ to the first register. This creates a uniform superposition over $G$: 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |x_0\rangle\,.
    \] 
    \item Now apply the group action in a controlled manner: for each basis state $|g\rangle$ in the first register, apply $g$’s action on the second register. This transforms the joint state into 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    which is an entangled state between a superposition of group elements and their corresponding $X$-elements.
    \item Apply the inverse QFT (which is the adjoint of the QFT) on the first register. This recombines the amplitudes in the first register. The result (by properties of the QFT) can be shown to be 
    \[
    \frac{1}{\sqrt{N}} \sum_{g \in G} \omega^{\langle h,\,g \rangle} |g\rangle \otimes |g * x_0\rangle\,,
    \] 
    where $\omega^{\langle h,\,g \rangle}$ denotes the phase picked up that depends on $h$ and $g$. In fact, this state can be written as $|h\rangle \otimes \frac{1}{\sqrt{N}}\sum_{g \in G} |g * x_0\rangle$; essentially, the first register ends up in $|h\rangle$ (the serial number) and the second register is an equal superposition of all elements of $X$ in the orbit of $x_0$ (which is $X$ itself, since the action is transitive) but weighted by a phase related to $h$. This is the so-called group-action Fourier state.
    \item The first register (which contains $|h\rangle$) is measured in the computational basis, yielding the random serial number $g$ and collapsing the second register to a state that we will denote $|\$_h\rangle$. This $|\$_h\rangle$ is the actual quantum money state corresponding to serial $h$.
    \item Output the pair $(s = h,\; |\$\rangle = |\$_h\rangle)$.
\end{enumerate}

@@

\section{Hartley-Transform Money Scheme Construction}
Our scheme modifies the above Fourier-based construction by utilizing the \textbf{quantum Hartley transform (QHT)} in place of the QFT. Since the Hartley transform is closely related to the Fourier transform (essentially it produces the real and imaginary parts in a single real-valued transform), one might expect this substitution to be straightforward. And indeed, for the generation procedure it largely is: we will use the QHT to create real-amplitude superposition states.

Concretely, assume again a group action $*: G\times X \mapsto X$ with $|G|=N$. We fix the same generating set $S$ and basepoint $x_0$. The $\mathsf{Gen}_{H}$ (Hartley generation) algorithm:
\begin{enumerate}
    \item Sample a random $h \leftarrow G$ to be the serial number.
    \item Prepare $|0_G, x_0\rangle$ as before.
    \item Apply the \emph{quantum Hartley transform} over $G$ to the first register. The Hartley transform (on an abelian group of order $N$) is defined as:
    \[
       \mathrm{QHT}: |g\rangle \mapsto \frac{1}{\sqrt{N}} \sum_{k \in G} \cas \Big(\frac{2\pi \langle k,g \rangle}{N}\Big) |k\rangle,
    \] 
    where $\mathrm{cas}(\theta) = \cos \theta + \sin \theta$ and $\langle k,g \rangle$ denotes some pairing like an exponent or dot product (for cyclic groups this is just $kg$). In simpler terms, $\mathrm{QHT}(|0_G\rangle) = \frac{1}{\sqrt{N}}\sum_{k \in G} |k\rangle$ (because $\cas(0)=1$), \emph{the same uniform superposition as the QFT produces for $|0\rangle$}. So after this step, we again have $\frac{1}{\sqrt{N}}\sum_{g \in G} |g\rangle|x_0\rangle$, but crucially the coefficients are all \emph{real} and positive (in fact all $1/\sqrt{N}$).
    \item Perform the controlled group action by the first register on the second: $\frac{1}{\sqrt{N}}\sum_{g} |g\rangle |g*x_0\rangle$.
    \item Apply the inverse QHT on the first register (which being its own inverse up to a constant factor, is the same operation again since Hartley is symmetric like Fourier). Now here is where things differ: after this operation, the joint state will be 
    \[
       \frac{1}{\sqrt{N}} \sum_{g \in G} \cas\Big(\frac{2\pi \langle h,g \rangle}{N}\Big)\, |g\rangle \otimes |g*x_0\rangle.
    \] 
    Because the Hartley transform combines what would be cosine and sine terms, this state is not as nicely factorable as in the Fourier case. In the Fourier case, we got $|h\rangle \otimes \text{(orbit state)}$. In the Hartley case, we do not get a single basis state in the first register; instead we get a superposition in the first register with real coefficients $c_g = \cas(2\pi\langle h,g\rangle/N)$. However, note that $c_g$ is just some real number in $[-1,1]$ depending on $h$ and $g$. Another way to express the above state is:
    \[
      |\$\rangle = \frac{1}{\sqrt{N}} \sum_{x \in X} \Big(\frac{1}{\sqrt{N}}\sum_{g: g*x_0 = x} \cas\big(\frac{2\pi \langle h,g \rangle}{N}\big)\Big)\, |x\rangle.
    \] 
    Since for a given $x\in X$, all $g$ such that $g*x_0=x$ have the form $g = u h$ for a unique $u \in G$ (because the action is free and transitive), one can simplify the internal sum using properties of the cas function. But we need not dive into that here.
    \item No measurement of the first register is performed. In fact, we do not have a nice product state separation, so we cannot simply measure one register to get $h$. Instead, the output of generation is just the quantum state $|\$\rangle$ (which is entangled between what used to be two registers) and the known classical serial number $h$ (we know what $h$ we picked).
\end{enumerate}

@@

% \begin{algorithm}[$\qht_N$] 
\paragraph{Algorithm $\qht_N$}
    \label{alg:qht-N}
    \begin{description}[font = \normalfont\itshape, itemsep = 0mm, parsep = 0mm, topsep = 1mm]
        \item [Input:] quantum state $\ket{\psi} \in \C^N$, where $N = 2^n$
        \item [Output:] quantum state $\qht_N\ket{\psi}$
    \end{description}

@@

\begin{theorem}
% \paragraph{Theorem.}
    \label{thm:qht-cost}
    Algorithm above is correct and can be implemented using $\approx \log^2 N + O(\log N)$ elementary gates.
\end{theorem}

\begin{proof}
% \paragraph{Proof.}
    The correctness of the algorithm follows from the preceding discussion. Except for the unitary \( U_R \) and the negation unitary in Step~\ref{stp:ngt}, all steps of the algorithm can be implemented using \( O(1) \) elementary gates. The negation operation in Step~\ref{stp:ngt} can be realized using approximately \( \lceil \log N \rceil \) elementary gates.

@@

    Let \( T(N) \) denote the gate complexity of the algorithm for an input of dimension \( N \). Assuming access to the \( R_j \) gates as elementary operations, the recurrence relation becomes
    \[
    T(N) \approx T(N / 2) + 2\log N + O(1).
    \]
    Solving this recurrence yields \( T(N) \approx \log^2 N + O(\log N) \), as claimed.
    \end{proof}
%%%%%%%%%%%  ta inja  %%%%%%%%%%%

@@

For the purposes of this thesis, the main point is: \emph{we can implement the Hartley transform needed for our money scheme with polynomial efficiency}. Thus, switching to the Hartley transform does not introduce any prohibitive cost. All steps of generation and verification remain efficient. Having covered the construction of the money scheme and the necessary tools, we now proceed to the final piece: the detailed verification algorithm using quantum walks, and an analysis of how it validates genuine banknotes and foils counterfeit attempts.
\vspace{1em}

@@

    \begin{lemma}
    % \paragraph{Lemma.}
        \label{lem:h-eigen}
        The money state $\ket{\mathbb{Z}_N^{(h)} * x}_H$ is an eigenstate of $W$ with eigenvalue $e^{i\lambda_h t}$.
    \end{lemma}
    \begin{proof}
    % \paragraph{Proof.}
        We have 
        \begin{align*}
            e^{iAt} \ket{\mathbb{Z}_N^{(h)} * x}_H
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x} \braket{\mathbb{Z}_N^{(g)} * x}{\mathbb{Z}_N^{(h)} * x}_H \\
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x} \bra{\mathbb{Z}_N^{(g)} * x} \Big( \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} \ket{\mathbb{Z}_N^{(-h)} * x} \Big) \\
            & = e^{i\lambda_{h}t} \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} e^{i\lambda_{-h}t} \ket{\mathbb{Z}_N^{(-h)} * x} \\
            & = e^{i\lambda_{h}t} \ket{\mathbb{Z}_N^{(h)} * x}_H,
        \end{align*}
        where the second equality follows from the identity in \eqref{eq:QHT_QFT}, and the last equality follows from the fact that $\lambda_h = \lambda_{-h}$.
    \end{proof}
